---
title: "Survival Predicted by Logistic Regression"
author: "Sarah VCH"
---
```{r}
library(randomForest)
library(Amelia)
library(dplyr)
library(ggplot2)
library(rmarkdown)
library(caTools)
library(stringr)
library(randomForest)

train <- read.csv("../input/train.csv")
test  <- read.csv("../input/test.csv")
```
# Did class affect survival?
#### To no ones surprise, it does looks like class affected survival. 
##### 63% of the first class passengers survived
##### 47% of the second class passengers survived
##### 24% of third class passengers survived



```{r}
#set Survied as factor 1 = Survived, 0 = Did not make it
train$Survived <- factor(train$Survived)
train$Pclass <- as.factor(train$Pclass)
levels(train$Pclass) <- c("1st Class", "2nd Class", "3rd Class")

fir.cl <-subset(train, train$Pclass == '1st Class')
sec.cl <-subset(train, train$Pclass == '2nd Class')
thr.cl <-subset(train, train$Pclass == '3rd Class')

#percent survival of first class
fir.perc <- (nrow(subset(fir.cl, fir.cl$Survived == 1))/nrow(fir.cl))*100

#percent survival of second class
sec.perc <- (nrow(subset(sec.cl, sec.cl$Survived == 1))/nrow(sec.cl))*100

#percent survival of first class
thr.per <- (nrow(subset(thr.cl, thr.cl$Survived == 1))/nrow(thr.cl))*100

class <- matrix(c(fir.perc,sec.perc,thr.per),ncol=1,byrow=FALSE)
colnames(class) <- c("Percent Survied")
rownames(class) <- c("1st", "2nd", "3rd")
class <- as.table(class)
class

ggplot(train,aes(Pclass)) + geom_bar(aes(fill=factor(Survived)),alpha=0.5, width =0.5, color='black') + scale_fill_discrete(name="Survived", breaks=c("0", "1"), labels=c("Died", "Survived")) + labs(title = "Survival Based on Class")
```

# Did gender affect survival?
#### Survival varied greatly with gender
##### 74% of female passengers survived
##### 19% of male passengers survived

```{r}
men <-subset(train, train$Sex == 'male')
women <-subset(train, train$Sex == 'female')
per.men <-(nrow(subset(men, men$Survived == 1))/nrow(men))*100
per.wom <-(nrow(subset(women, women$Survived == 1))/nrow(women))*100

gender <- matrix(c(per.men, per.wom),ncol=1,byrow=FALSE)
colnames(gender) <- c("Percent Survied")
rownames(gender) <- c("men","women")
gender <- as.table(gender)
gender

ggplot(train,aes(Sex)) + geom_bar(aes(fill=factor(Survived)), alpha=0.5, color='black') + scale_fill_discrete(name="Survived", breaks=c("0", "1"), labels=c("Died", "Survived")) + labs(title = "Survival Based on Gender")
```

# Did age affect Survival?
#### It looks like age had some impact on passenger survival
##### 58% of passengers under 15 survied
##### 38% of passengers 15-33 survied
##### 42% of passengers 34-55 survied
##### 31% of passengers over 56 survived

```{r}
kids <- subset(train, train$Age <= 14)
young <- subset(train, train$Age >= 15 & train$Age <= 33)
midage <- subset(train, train$Age >=34 & train$Age <= 55)
older <- subset(train, train$Age >= 56)

kids.sur <-nrow(subset(kids, kids$Survived == 1))/nrow(kids)*100
young.sur <-nrow(subset(young, young$Survived == 1))/nrow(young)*100
midage.sur <-nrow(subset(midage, midage$Survived == 1))/nrow(midage)*100
older.sur <- nrow(subset(older, older$Survived == 1))/nrow(older)*100

age1 <- matrix(c(kids.sur,young.sur,midage.sur,older.sur),ncol=1,byrow=FALSE)
colnames(age1) <- c("Percent Survied")
rownames(age1) <- c("under 15","15-33","34-55","56+")
age1 <- as.table(age1)
age1


ggplot(train,aes(Age)) + geom_histogram(aes(fill=factor(Survived)), color='black', alpha=0.5, binwidth =5) + scale_fill_discrete(name="Survived", breaks=c("0", "1"), labels=c("Died", "Survived")) + labs(title = "Survival Based on Age")
```

## Did fare effect Survival?
#### It does look like passenger survival increased with fare
##### For the 651 passengers that paid under $30 only 30% Survived
##### For the 79 passengers that paid between $30-$50 41% Survived
##### For the 161 passengers that paid over $50 68% Survived

```{r}
under.30 <-subset(train, train$Fare < 30)
bt30.50 <-subset(train, train$Fare >= 30 & train$Fare < 50)
over.50 <-subset(train, train$Fare >= 50)

nrow(under.30)
nrow(bt30.50)
nrow(over.50)


under <- nrow(subset(under.30, under.30$Survived ==1))/(nrow(under.30))*100
btw <-  nrow(subset(bt30.50, bt30.50$Survived ==1))/(nrow(bt30.50))*100
over <- nrow(subset(over.50, over.50$Survived ==1))/(nrow(over.50))*100

fr <- matrix(c(under,btw,over),ncol=1,byrow=FALSE)
colnames(fr) <- c("Percent Survied")
rownames(fr) <- c("under 30","30-50","over 50")
fr <- as.table(fr)
fr

ggplot(train,aes(Fare)) + geom_histogram(aes(fill=factor(Survived)), color='black', alpha=0.5) + scale_fill_discrete(name="Survived", breaks=c("0", "1"), labels=c("Died", "Survived")) + labs(title = "Survival Based on Fare")
```

# Did it matter where people boarded?
#### C = Cherbourg, Q = Queenstown, S = Southampton
#### 55% of passengers that boarded from Cherbourg survived
#### 39%% of passengers that boarded from Queenstown survived
#### 34% of passengers that boarded from Southhampton survived

# Did more men board in Southhampton?
#### In Southhampton 68% of the passengers were men where as Cherbour and Qeenstown it was closer to 50% 

```{r}

c <- subset(train, train$Embarked == "C")
q <- subset(train, train$Embarked == "Q")
s <- subset(train, train$Embarked == "S")

c.per <- nrow(subset(c, c$Survived == 1))/(nrow(c))*100
s.per <- nrow(subset(s, s$Survived == 1))/(nrow(s))*100
q.per <- nrow(subset(q, q$Survived == 1))/(nrow(q))*100

loc <- matrix(c(c.per,q.per, s.per),ncol=1,byrow=FALSE)
colnames(loc) <- c("Percent Survied")
rownames(loc) <- c("Cherbourg", "Queenstown", "Southhampton")
loc <- as.table(loc)
loc

c.men <- nrow(subset(c, c$Sex == 'male'))/(nrow(c))*100
s.men <- nrow(subset(s, s$Sex == 'male'))/(nrow(s))*100
q.men <- nrow(subset(q, q$Sex == 'male'))/(nrow(q))*100

sx <- matrix(c(c.men,q.men, s.men),ncol=1,byrow=FALSE)
colnames(sx) <- c("Percent Men")
rownames(sx) <- c("Cherbourg", "Queenstown", "Southhampton")
sx <- as.table(sx)
sx

ggplot(train,aes(Embarked)) + geom_bar(aes(fill=factor(Survived)), color='black', alpha=0.5) + scale_fill_discrete(name="Survived", breaks=c("0", "1"), labels=c("Died", "Survived")) + labs(title = "Survival Based on Boarding Location")
ggplot(train,aes(Embarked, Fare)) + geom_point(aes(color=factor(Sex)), alpha=0.5, position='jitter') + coord_cartesian(ylim=c(0, 300))
```

#Lets remove the blank Embarked rows

```{r}

train$Embarked <- as.character(train$Embarked)
train <- subset(train, train$Embarked != "")
train$Embarked <- as.factor(train$Embarked)
str(train)

```

# Did passengerId affect Survival?
#### Looking at the graph below it doesn not appear that Passenger Id affected survival

```{r}
ggplot(train,aes(Survived, PassengerId)) + geom_boxplot(aes(color=factor(Survived))) + labs(title ='Survival based on Id')
```

# Did the number of Siblings and Spouses effect Survival?
#### SibSp = # of siblings / spouses aboard the Titanic
#### Passengers faired better with one or two other significant others
Only 34% of passengers traveling alone survived. A much higher percent of passengers survived when traveling with another. Passengers traveling with another had a 54% survival. 
Passengers traveling with two others had a 46% survival. Passengers traveling with more than two other sibling or spouses had a lower percent of survival. Passengers with three siblings or spouses had a 25% change of survival. 
Passengers traveling with four sibling or spouses had a 16% of survival. Passengers with over four sibliings or spouses did not survive


```{r}
zero <- subset(train, train$SibSp == 0)
one <- subset(train, train$SibSp == 1)
two <- subset(train, train$SibSp == 2)
three <- subset(train, train$SibSp == 3)
four <- subset(train, train$SibSp == 4)
five <- subset(train, train$SibSp == 5)
six <- subset(train, train$SibSp == 6)
seven <- subset(train, train$SibSp == 7)
eight <- subset(train, train$SibSp == 8)

zero.per <- nrow(subset(zero, zero$Survived == 1))/nrow(zero)*100
one.per <- nrow(subset(one, one$Survived == 1))/(nrow(one))*100
two.per <- nrow(subset(two, two$Survived == 1))/(nrow(two))*100
three.per <- nrow(subset(three, three$Survived == 1))/(nrow(three))*100
four.per <- nrow(subset(four, four$Survived == 1))/(nrow(four))*100
five.per <- nrow(subset(five, five$Survived == 1))/(nrow(five))*100
six.per <- nrow(subset(six, six$Survived == 1))/(nrow(six))*100
seven.per <- nrow(subset(seven, seven$Survived == 1))/(nrow(seven))*100
eight.per <- nrow(subset(eight, eight$Survived == 1))/(nrow(eight))*100

ss <- matrix(c(zero.per, one.per, two.per, three.per, four.per, five.per, six.per, seven.per, eight.per,nrow(zero),nrow(one), nrow(two),nrow(three), nrow(four), nrow(five), nrow(six), nrow(seven), nrow(eight)),ncol=2,byrow=FALSE)
colnames(ss) <- c("Percent Survived", "Count in Group")
rownames(ss) <- c("0","1", "2", "3", "4","5","6","7","8")
ss <- as.table(ss)
ss


ggplot(train,aes(SibSp)) + geom_histogram(aes(fill=factor(Survived)), color='black', alpha=0.5) + scale_fill_discrete(name="Survived", breaks=c("0", "1"), labels=c("Died", "Survived")) + labs(title = "Survival Based on # Siblings and Spouses")
```

# Did the number of Parents and Children affect survival?
#### Again it looks like people faired best with the addition of one other
#### 34% of Passengers traveling without parent or child survived
#### 55% of Passengers traveling with another survived
#### 50% of Passengers traveling with two parents and or children survived
#### 27% of Passengers traveling with three or more parents and or children survived

```{r}

par.zero <- subset(train, train$Parch == 0)
par.one <- subset(train, train$Parch == 1)
par.two <- subset(train, train$Parch == 2)
par.threeup <- subset(train, train$Parch >=3)

per.zero <- nrow(subset(par.zero, par.zero$Survived == 1))/nrow(par.zero)*100
per.one <- nrow(subset(par.one, par.one$Survived == 1))/nrow(par.one)*100
per.two <- nrow(subset(par.two, par.two$Survived == 1))/nrow(par.two)*100
per.threeup <- nrow(subset(par.threeup, par.threeup$Survived == 1))/nrow(par.threeup)*100

pc <- matrix(c(per.zero, per.one, per.two, per.threeup, nrow(par.zero), nrow(par.two), nrow(par.threeup)),ncol=2,byrow=FALSE)
colnames(pc) <- c("Pecent Survival", "Count in Group")
rownames(pc) <- c("0", "1", "2", "3+")
pc <- as.table(pc)
pc

table(train$Parch)

change <-function(parchi){
  for(i in 1:length(parchi)){
    if(parchi[i] == 9 | parchi[i] == 6){
      parchi[i] <- 5
    }else{
      parchi[i] <- parchi[i]
    }
  }
  return(parchi)
}

train$Parch <-change(train$Parch)

ggplot(train,aes(Parch)) + geom_bar(aes(fill=factor(Survived)), color='black', alpha=0.5) + scale_fill_discrete(name="Survived", breaks=c("0", "1"), labels=c("Died", "Survived")) + labs(title = "Survival Based on Number of Parents and Children")

```

# Can we gather information based on cabin?
#### It looks like there is mess of information in the cabin column and many missing values
#### We could add all Cabin information to the model
#### We could also simplify the Cabin information down to Known and Unknown


```{r}
table(train$cabin)

train$Cabin <- as.character(train$Cabin)

splt <- function(cabin){
  for(i in 1:length(cabin)){
    if(cabin[i] == ''){
      cabin[i] <- cabin[i]
    }else{
      cabin[i] <- substring(cabin[i],1,1)
    }
  }
  return(cabin)
}

train$Cabin <- splt(train$Cabin)
ggplot(train, aes(Cabin)) + geom_bar(aes(fill=factor(Survived)),color='black', alpha=0.5) + scale_fill_discrete(name="Survived", breaks=c("0", "1"), labels=c("Died", "Survived")) + labs(title = "Survival Based on Cabin")

known <- function(cabin){
  for(i in 1:length(cabin)){
    if(cabin[i] == ''){
      cabin[i] <- 'unknown'
    }else{
      cabin[i] <- 'known'
    }
  }
  return(cabin)
}

train$Cabin <- known(train$Cabin)

ggplot(train, aes(Cabin)) + geom_bar(aes(fill=factor(Survived)),color='black', alpha=0.5) + scale_fill_discrete(name="Survived", breaks=c("0", "1"), labels=c("Died", "Survived")) + labs(title = "Cabin Known vs Unknown")


```

# Can we gather any information based on name?
#### It looks like again women seemed to fair better than men
#### I've condensed a few things down:


```{r}
s <- data.frame(str_split_fixed(train$Name, ",", 2))
s1 <- data.frame(str_split_fixed(s$X2, ". ", 2))
train$Name <- s1$X1
str(train$Name)
table(train$Name)
ggplot(train, aes(factor(Name))) + geom_bar(aes(fill=factor(Survived)),color='black', alpha=0.5) + labs(title ='Title')
train$Name <- as.character(train$Name)

name.change <- function(title){
  for(i in 1:length(title)){
    if(title[i] == ' Master' ){
      title[i] <- ' Mr'
    }else if (title[i] == ' Ms'| title[i] == ' Lady' | title[i] == ' Dona'| title[i] == ' Mme' |title[i] == ' Mlle' | title[i] == ' th' |title[i] == ' Mrs' |title[i] == ' Miss'){
      title[i] <- ' Ms'
    }else if(title[i] == ' Don' | title[i] == ' Sir' | title[i] == ' Jonkheer' | title[i] == ' Col' | title[i] == ' Major' |title[i] == ' Capt' | title[i] == ' Dr'){
      title[i] <- ' Titled Men'
    }else{
      title[i] <- title[i]
    }
  }
  return(title)
}

train$Name <- sapply(train$Name, name.change)
train$Name <- as.factor(train$Name)
table(train$Name)

ggplot(train, aes(factor(Name))) + geom_bar(aes(fill=factor(Survived)), color='black', alpha=0.5) + labs(title ='Title')

```

# What information went down with the ship?
```{r}
missmap(train, main="Titanic Training Data - Missing Map", col=c('yellow', 'black'), y.labels=c(''), y.at =c(1), legend=FALSE)
```

# Does age vary by class?
#### first class has the highest mean age followed by second class and third class
#### I'm filling in missing passengers ages with the average class age

```{r}
ggplot(train, aes(factor(Pclass), Age)) + geom_boxplot(aes(color=factor(Pclass))) + labs(title ='Age of Passengers by Class')

train$Pclass <- as.character(train$Pclass)
change.class <-function(classn){
  for(i in 1:length(classn)){
    if(classn[i] == '1st Class'){
      classn[i] <- 1
    }else if(classn[i] == '2nd Class'){
      classn[i] <- 2
    }else if(classn[i] == '3rd Class'){
      classn[i] <- 3
    }
  }
  return(classn)
}

train$Pclass <-change.class(train$Pclass)
str(train$Pclass)

mean((subset(train, train$Pclass == 1))$Age, na.rm=TRUE)
mean((subset(train, train$Pclass == 2))$Age, na.rm=TRUE)
mean((subset(train, train$Pclass == 3))$Age, na.rm=TRUE)
```

# Replace Age NA values with mean age from class
#### first class, 38.23
#### second class, 29.88
#### thrid class, 25.14

```{r}
fillin_age <- function(age, class){
  out <- age
  for(i in 1:length(age)){
    if(is.na(age[i])){
      if(class[i] == 1){
        out[i] <-38.23
      }else if(class[i] ==2){
        out[i] <-29.88
      }else{
        out[i] <-25.14
      }
    }else{
      out[i] <-age[i]
    }
  }
  return(out)
}

train$Age <-fillin_age(train$Age, train$Pclass)
```

#Check for missing values in Age
#### Looks like age is filled up nicely with new imputed age

```{r}
missmap(train, main='Missing data', col=c("yellow","black"), legend = FALSE)
```

# Update feilds to factors and take out PassengerId and Ticket

```{r}
train$Survived <- as.factor(train$Survived)
train$Pclass <- as.factor(train$Pclass)
train$SibSp <- as.factor(train$SibSp)
train$Parch <- as.factor(train$Parch)
train$Cabin <- as.factor(train$Cabin)
train <- select(train,-PassengerId,-Ticket)
```

# Train the model with subst of train data
#### model = Survied preditced based off of all columns except Passenger ID and Ticket
#### lowest P values: Pclass = 3, Age, SibSp = 3, Cabin = Unknown

```{r}
split = sample.split(train$Survived, SplitRatio = 0.70)
final.train = subset(train, split == TRUE)
final.test = subset(train, split == FALSE)

final.log.model <- glm(formula=Survived ~ . , family = binomial(link='logit'),data = final.train)
summary(final.log.model)
fitted.probabilities <- predict(final.log.model,newdata=final.test,type='response')
```

# Confusion Matrix set up and Accuracy calculated

```{r}
d <-table(final.test$Survived, fitted.probabilities > 0.5)
d
sum(diag(d))/sum(d)
```

# Check and clean test data
#### Impute age
#### transform Pclass
#### Change Parch and SibSp to Factors
#### Change Cabin to known and unknown
```{r}
ggplot(test,aes(Sex)) + geom_bar(color='black',alpha=0.5) + labs(title ='Count of Men and Women Aboard')

#change NA values to 0
missmap(test, main='Missing data', col=c("yellow","black"), legend = FALSE)
summary(test$Fare)
test$Fare[is.na(test$Fare)] <- 0
missmap(test, main='Missing data', col=c("yellow","black"), legend = FALSE)

#Name
s <- data.frame(str_split_fixed(test$Name, ",", 2))
s1 <- data.frame(str_split_fixed(s$X2, ". ", 2))
test$Name <- s1$X1
str(test$Name)
table(test$Name)
test$Name <- as.character(test$Name)
test$Name <- sapply(test$Name, name.change)


#Age
test$Age <-fillin_age(test$Age, test$Pclass)

#Parent Children
test$Parch <-change(test$Parch)

#Cabin
test$Cabin <- as.character(test$Cabin)
test$Cabin <- splt(test$Cabin)
test$Cabin <- known(test$Cabin)
test$Cabin <- as.factor(test$Cabin)


#Factor Things Up
test$Pclass <- as.factor(test$Pclass)
test$Parch <- as.factor(test$Parch)
test$SibSp <- as.factor(test$SibSp)
test$Name <- as.factor(test$Name)

test <- select(test,-PassengerId,-Ticket)
```

# Use model to predict Survival

```{r}
test$fitted.probabilities = predict(final.log.model, newdata=test, type="response")
fitted.results <- ifelse(test$fitted.probabilities > 0.5,1,0)
test$Survived <-fitted.results
```

#Percent Survival

```{r}
sum.surv <- subset(test, test$Survived == 1)
perc.surv <-nrow(sum.surv)/nrow(test)
```

#predicted precent of survial on test data

```{r}
print(perc.surv)
```

#actual percent of survival for the titanic

```{r}
print((2224-1502)/2224)
```

#Check out old data visually for accuracy of prediction

```{r}
test$Survived <- as.factor(test$Survived)
ggplot(test,aes(Sex)) + geom_bar(aes(fill=factor(Survived)), alpha=0.5) + labs(title ='Count of Men and Women Aboard')
ggplot(test, aes(factor(Name))) + geom_bar(aes(fill=factor(Survived)), alpha=0.5) + labs(title ='Title')

```

# Try Random Forest to see if accuracy improves
#### Similar accuracy to Logistic Regression


```{r}
rf.model <- randomForest(Survived ~., data=final.train)
predicted.values1 <- predict(rf.model,final.test[2:10])
d <- table(predicted.values1,final.test$Survived)
print(d)
sum(diag(d))/sum(d)

```
# Use all train data for new Logistic Regression model
```{r}

final.final.log.model <- glm(Survived ~ . , family = binomial(link='logit'),data = train)
summary(final.final.log.model)

fitted.probabilities1 <- predict(final.final.log.model,newdata=test[1:9],type='response')
test$fitprob <- fitted.probabilities1
fitted.results1 <- ifelse(test$fitprob > 0.5,1,0)

test$Survived1 <-fitted.results1
sum.surv <- subset(test, test$Survived1 == 1)
perc.surv <-nrow(sum.surv)/nrow(test)
print(perc.surv)
```
# Actual percent of passengers that survived on the Titanic

```{r}
print((2224-1502)/2224)
```


#### Check out old data visually for accuracy of prediction

```{r}
ggplot(test,aes(Sex)) + geom_bar(aes(fill=factor(Survived1)), alpha=0.5) + labs(title ='Count of Men and Women Aboard')
ggplot(test, aes(factor(Name))) + geom_bar(aes(fill=factor(Survived1)), alpha=0.5) + labs(title ='Title')

```

# Write the solution to file

```{r}

test1  <- read.csv("../input/test.csv")

submission <- data.frame(PassengerID = test1$PassengerId, Survived = test$Survived)
head(submission)
write.csv(submission, file = 'rf_predictions.csv', row.names = F)






